---
format:
  html: 
    warning: false 
    echo: false 
---

## Exploratory Data Analysis

This is my preliminary EDA. It's what I presented to the doctor (and I will update this) so it's not on every variable and not as extensive as the gtsummary on the data summary page. 

### Initial Data Summary 

```{r}
library(tidyverse)
library(dplyr)
library(gt)

df <- read_csv("../data/cleaned_data/cleaned_data.csv", col_names = TRUE)
cols_drop <- c("...1")
df <- df %>% dplyr::select(-all_of(cols_drop))
df <- df %>% filter(exclude == 0)

df <- df %>% filter(!is.na(patient_id))
df <- df %>% filter(!is.na(gender))
```

#### Gender 

```{r}
gender_df <- data.frame(table(df$gender))
colnames(gender_df) <- c("Gender", "Count")
gender_df %>% gt()
```

#### Diagnosis Age

```{r}
diag_age_summary <- summary(df$diagnosis_age)
diag_age_df <- data.frame(Statistic = names(diag_age_summary),
                          Value = as.numeric(diag_age_summary))
diag_age_df %>% gt()

ggplot(df, aes(x = diagnosis_age)) +
  geom_density(fill = "darkblue", alpha = 0.5) +
  labs(x = "Age", y = "Density", title = "Age Density Plot") +
  theme(plot.title = element_text(hjust = 0.5))
```
#### Race

```{r}

race_sum <- table(df$race)
race_df <- data.frame(Race = names(race_sum),
                    Count = as.numeric(race_sum))
race_df %>% gt()

ggplot(df, aes(x = race, fill = race)) + 
  geom_bar(fill = "darkblue") + 
  labs(x = "Race", y = "Count", title = "Count by Race in the Data") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")
```


#### Vital Status

```{r}
vs_df_sum <- table(df$vital_status)
vs_df <- data.frame(Status = names(vs_df_sum),
                    Count = as.numeric(vs_df_sum))
vs_df %>% gt()

ggplot(df, aes(x = vital_status, fill = vital_status)) + 
  geom_bar(fill = "darkblue") + 
  labs(x = "Vital Status", y = "Count", title = "Count by Vital Status in the Data") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")
```

#### Treatment Summary 

##### Radiation Treatment 

```{r}
df$num_radiations <- factor(df$num_radiations, levels = c("0", "<1", "1", "<2", "2", "≥3"))
rt_df_sum <- table(df$num_radiations)
rt_df <- data.frame(Cycles = names(rt_df_sum),
                    Count = as.numeric(rt_df_sum))
rt_df$Percentage <- round((rt_df$Count/sum(rt_df$Count)*100), 2)
rt_df %>% gt()
```

##### Chemotherapy Treatment 

```{r}
df$num_chemo <- factor(df$num_chemo, levels = c("0", "<1", "1", "<2", "2", "≥3"))
ct_df_sum <- table(df$num_chemo)
ct_df <- data.frame(Cycles = names(ct_df_sum),
                    Count = as.numeric(ct_df_sum))
ct_df$Percentage <- round((ct_df$Count/sum(ct_df$Count)*100), 2)
ct_df %>% gt()
```

##### Stupp Protocol 

```{r}
stupp_df_sum <- table(df$received_stupp)
stupp_df <- data.frame(Cycles = names(stupp_df_sum),
                    Count = as.numeric(stupp_df_sum))
stupp_df %>% gt()
```

#### Time from Surgery to Start of Radiation 

```{r}
df_surgrad <- df %>% filter(!is.na(df$time_surgery_radiation))
surgrad_df_sum <- summary(df_surgrad$time_surgery_radiation)
surgrad_df <- data.frame(Statistics = names(surgrad_df_sum),
                    Value = as.numeric(surgrad_df_sum))
surgrad_df %>% gt()
```

#### Number of Lobes Involved

```{r}
df_lobes <- df %>% filter(!is.na(df$num_lobes))
lobes_df_sum <- table(df_lobes$num_lobes)
lobes_df <- data.frame(Lobes = names(lobes_df_sum),
                    Count = as.numeric(lobes_df_sum))
lobes_df %>% gt()
```

### Path Data

#### Microvascular Proliferation

```{r}
library(rlang)

summarize_variable <- function(data, var) {
  var_sym <- ensym(var)
  var_label <- paste0(as_label(var_sym))
  
  data %>%
    filter(!is.na(!!var_sym)) %>%
    count(!!var_sym, name = "Count") %>%
    rename(!!var_label := !!var_sym) %>%
    gt()
}

summarize_variable(df, microvascular_proliferation)
```

#### Necrosis

```{r}
summarize_variable(df, necrosis)
```

#### Caris?

```{r}
summarize_variable(df, caris)
```

#### EGFR Amplification

```{r}
summarize_variable(df, EGFR_amplification)
```

#### +7/-10 Chromosomal Alterations

```{r}
summarize_variable(df, chromosomal710)
```

#### TERT Promotor

```{r}
summarize_variable(df, TERT_promoter)
```

#### TERT Promotor

```{r}
summarize_variable(df, pTERT_variant)
```

#### MGMT Methylation

```{r}
summarize_variable(df, MGMT_methylation)
```

#### Ki67

```{r}
summarize_variable(df, Ki67_LI)
```

#### ATRX

```{r}
summarize_variable(df, ATRX)
```

#### Mitotic Activity

```{r}
summarize_variable(df, mitotic_activity_increase)

```
