---
format:
  html: 
    warning: false 
    echo: false 
---

### Initial Data Summary 

```{r}
library(tidyverse)
library(dplyr)
library(gt)

df <- read_csv("../data/cleaned_data/cleaned_data.csv", col_names = TRUE)
df <- df %>% filter(exclude == 0)
```

#### Gender 

```{r}
gender_df <- data.frame(table(df$gender))
colnames(gender_df) <- c("Gender", "Count")
gender_df %>% gt()
```

#### Diagnosis Age

```{r}
diag_age_summary <- summary(df$diagnosis_age)
diag_age_df <- data.frame(Statistic = names(diag_age_summary),
                          Value = as.numeric(diag_age_summary))
diag_age_df %>% gt()

ggplot(df, aes(x = diagnosis_age)) +
  geom_density(fill = "darkblue", alpha = 0.5) +
  labs(x = "Age", y = "Density", title = "Age Density Plot") +
  theme(plot.title = element_text(hjust = 0.5))
```
#### Race

```{r}
ggplot(df, aes(x = race, fill = race)) + 
  geom_bar(fill = "darkblue") + 
  labs(x = "Race", y = "Count", title = "Count by Race in the Data") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")
```


#### Vital Status

```{r}
vs_df_sum <- table(df$vital_status)
vs_df <- data.frame(Status = names(vs_df_sum),
                    Count = as.numeric(vs_df_sum))
vs_df %>% gt()

ggplot(df, aes(x = vital_status, fill = vital_status)) + 
  geom_bar(fill = "darkblue") + 
  labs(x = "Vital Status", y = "Count", title = "Count by Vital Status in the Data") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")
```

#### Treatment Summary 

##### Radiation Treatment 

```{r}
df$num_radiations <- factor(df$num_radiations, levels = c("0", "<1", "1", "<2", "2", "≥3"))
rt_df_sum <- table(df$num_radiations)
rt_df <- data.frame(Cycles = names(rt_df_sum),
                    Count = as.numeric(rt_df_sum))
rt_df$Percentage <- round((rt_df$Count/sum(rt_df$Count)*100), 2)
rt_df %>% gt()
```

##### Chemotherapy Treatment 

```{r}
df$num_chemo <- factor(df$num_chemo, levels = c("0", "<1", "1", "<2", "2", "≥3"))
ct_df_sum <- table(df$num_chemo)
ct_df <- data.frame(Cycles = names(ct_df_sum),
                    Count = as.numeric(ct_df_sum))
ct_df$Percentage <- round((ct_df$Count/sum(ct_df$Count)*100), 2)
ct_df %>% gt()
```

##### Stupp Protocol 

```{r}
stupp_df_sum <- table(df$received_stupp)
stupp_df <- data.frame(Cycles = names(stupp_df_sum),
                    Count = as.numeric(stupp_df_sum))
stupp_df %>% gt()
```

#### Time from Surgery to Start of Radiation 

```{r}
df_surgrad <- df %>% filter(!is.na(df$time_surgery_radiation))
surgrad_df_sum <- summary(df_surgrad$time_surgery_radiation)
surgrad_df <- data.frame(Statistics = names(surgrad_df_sum),
                    Value = as.numeric(surgrad_df_sum))
surgrad_df %>% gt()
```

#### Number of Lobes Involved

```{r}
df_lobes <- df %>% filter(!is.na(df$num_lobes))
lobes_df_sum <- table(df_lobes$num_lobes)
lobes_df <- data.frame(Lobes = names(lobes_df_sum),
                    Count = as.numeric(lobes_df_sum))
lobes_df %>% gt()
```

### Path Data

#### Microvascular Proliferation

```{r}
library(rlang)

summarize_variable <- function(data, var) {
  var_sym <- ensym(var)
  var_label <- paste0(as_label(var_sym))
  
  data %>%
    filter(!is.na(!!var_sym)) %>%
    count(!!var_sym, name = "Count") %>%
    rename(!!var_label := !!var_sym) %>%
    gt()
}

summarize_variable(df, microvascular_proliferation)
```

#### Necrosis

```{r}
summarize_variable(df, necrosis)
```

#### Caris?

```{r}
summarize_variable(df, caris)
```

#### EGFR Amplification

```{r}
summarize_variable(df, EGFR_amplification)
```

#### +7/-10 Chromosomal Alterations

```{r}
summarize_variable(df, chromosomal710)
```

#### TERT Promotor

```{r}
summarize_variable(df, TERT_promoter)
```

#### TERT Promotor

```{r}
summarize_variable(df, pTERT_variant)
```

#### MGMT Methylation

```{r}
summarize_variable(df, MGMT_methylation)
```

#### Ki67

```{r}
summarize_variable(df, Ki67_LI)
```

#### ATRX

```{r}
summarize_variable(df, ATRX)
```

#### Mitotic Activity

```{r}
summarize_variable(df, mitotic_activity_increase)
```

### Initial Analysis 

#### Overall Survival 

This is a general picture of the overall survival of everyone in the data set. The red line represents the median value. 

```{r}
library(survival)

km <- survfit(Surv(time = df$overall_survival, event = df$status) ~ 1, data = df)
plot(km, xlab = "Time", ylab = "Survival Probability", main = "Kaplan-Meier Curve", conf.int = FALSE)
abline(v = km$time[km$surv <= 0.5][1], col = "red", lty = 2)
```

### Significant Variables 

Below are all of the individually significant variables I found in the dataset. 

```{r}
library(ggsurvfit)
df$age_group <- cut(df$diagnosis_age,
                    breaks = c(30, 40, 50, 60, 70, 80, 90, 100),  
                    labels = c("30-39", "40-49", "50-59", "60-69", "70-79", "80-89", "90-99"),  
                    right = FALSE)

sig_results <- data.frame(variable = character(), p_value = numeric())

categorical_vars <- c(
  "gender", "race", "institution_type", "resection_extent",
  "num_surgeries", "num_radiations", "num_chemo", "discontinued_treatment",
  "received_stupp", "discontinued_stupps", "on_treatment_last_visit", "lobes_involved",
  "other_brain_regions", "unique_clinical_findings", "seizures_presentation",
  "seizures_later", "dvt_pe_presentation", "dvt_pe_during", "microvascular_proliferation",
  "mp_early", "mp_focal", "necrosis", "chromosomal710", "EGFR_amplification",
  "TERT_promoter", "pTERT_variant", "MGMT_methylation", "Ki67_LI", "ATRX",
  "other_mutations", "mitotic_activity_increase", "specific_mitotic", "caris",
  "dna_profile", "methylation_result", "lobes_clean", "frontal", "parietal",
  "temporal", "occipital", "age_group"
)

for (variable in categorical_vars) {
  formula <- as.formula(paste("Surv(overall_survival, status == 1) ~", variable))
  surv_diff <- survdiff(formula, data = df)
  p <- 1 - pchisq(surv_diff$chisq, df = length(surv_diff$n) - 1)
  
  if (!is.na(p) && p < 0.05) {
    sig_results <- rbind(sig_results, data.frame(variable = variable, p_value = p))
  }
}

sig_results <- sig_results[order(sig_results$p_value), ]

sig_variables <- c("num_radiations", "num_chemo", "received_stupp", "discontinued_stupps", "num_surgeries",
                   "resection_extent", "discontinued_treatment", "caris", "MGMT_methylation", "temporal")
sig_results <- sig_results %>% filter(variable %in% sig_variables)
sig_results %>% gt()
```

### K-M Curves for Sig Variables 

```{r}
df$num_radiations <- factor(df$num_radiations, levels = c("0", "<1", "1", "<2", "2", "≥3"))
df$num_chemo <- factor(df$num_chemo, levels = c("0", "<1", "1", "<2", "2", "≥3"))
df$num_surgeries <- factor(df$num_surgeries, levels = c("1", "2", "≥3"))
df$resection_extent <- factor(df$resection_extent, levels = c("Biopsy", "Resection NOS", "Subtotal Resection", "Gross Total Resection"))

for (variable in sig_variables) {
  formula <- as.formula(paste("Surv(overall_survival, status == 1) ~", variable))
  
  plot <- survfit2(formula, data = df) |>  
    ggsurvfit() +
    add_pvalue("annotation", size = 5) +
    labs(
      x = "Time",
      title = paste("Survival Analysis of", variable)
    )
  
  print(plot)
}
```

### Next Steps

The next steps for the analysis will be: \
- Work on cleaning, categorising, and analysing the other mutations, increased mitotic activity values, and lobes involved. \
- Looking at how these variables may interact and influence each other and the corresponding survival time. \
- Meeting with my professor next week to discuss any avenues of analysis I may be missing to ensure that all of the data work is strong. 
