---
format:
  html: 
    warning: false 
    echo: false 
---

```{r}
library(tidyverse)
library(dplyr)
library(gt)
library(survival)
library(survminer)
library(dplyr)
library(ggsurvfit)

df <- read_csv("../data/cleaned_data/cleaned_data.csv")
cols_drop <- c("...1")
df <- df %>% dplyr::select(-all_of(cols_drop))
df <- df %>% filter(exclude == 0)
```

### Treatment Summary

```{r}
stupps <- mean(df$received_stupp == "yes") * 100
chemorad_ns <- df %>% filter(num_chemo != "0" & num_radiations != "0" & received_stupp == "no") %>% summarise(pct = n() / nrow(df) * 100) %>% pull(pct)
only_rt <- df %>% filter(num_chemo == "0" & num_radiations != "0") %>% summarise(pct = n() / nrow(df) * 100) %>% pull(pct)
only_chemo <- df %>% filter(num_chemo != "0" & num_radiations == "0") %>% summarise(pct = n() / nrow(df) * 100) %>% pull(pct)
no_treat <- df %>% filter(num_chemo == "0" & num_radiations == "0") %>% summarise(pct = n() / nrow(df) * 100) %>% pull(pct)

treatment_summary <- data.frame(
  treatment = c("Stupp Protocol", "Chemo/Radiation No Stupp", "Only RT", "Only Chemo", "No Treatment"),
  percent = c(stupps, chemorad_ns, only_rt, only_chemo, no_treat)
)

colnames(treatment_summary) <- c("Treatment Type", "Percent")
treatment_summary %>% gt()
```

### Is there a difference in overall survival for MGMT unmethylated and methylated status patients who received the Stupp protocol? 

```{r}
df_stupps <- df %>% filter(received_stupp == "yes")

km <- survfit2(Surv(overall_survival, status == 1) ~ MGMT_methylation, data = df_stupps)
ggsurvfit(km) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival of Patients Treated with Stupps by MGMT Status"
    )
```


```{r}
df$stupps_mgmt <- interaction(df$received_stupp, df$MGMT_methylation)

fit <- survfit(Surv(overall_survival, status == 1) ~ stupps_mgmt, data = df)
medians <- surv_median(fit)

medians_df <- data.frame(group = names(summary(fit)$table[, "median"]),
                         median_survival = summary(fit)$table[, "median"])

medians_df <- medians_df %>%
  separate(group, into = c("received_stupp", "mgmt_status"), sep = "\\.") %>%
  mutate(received_stupp = sub(".*=", "", received_stupp))  

colnames(medians_df) <- c("Received Stupps?", "MGMT Status", "Median Survival (Months)")

medians_df %>% gt()
```

We see that Stupps + Methylated Status = longest median survival time. Kruskal-Wallis test to test for significance.

```{r}
df_stupps <- df %>% filter(stupps_mgmt %in% c("no.M", "yes.M", "no.UM", "yes.UM"))

ggboxplot(df_stupps, 
          x = "stupps_mgmt", 
          y = "overall_survival", 
          color = "stupps_mgmt", 
          order = c("no.M", "yes.M", "no.UM", "yes.UM"),
          ylab = "Overall Survival", 
          xlab = "Stupps/MGMT Combination") +
  scale_x_discrete(labels = c("no.M" = "No Stupp\nMethylated", 
                              "yes.M" = "Yes Stupp\nMethylated",
                              "no.UM" = "No Stupp\nUnmethylated", 
                              "yes.UM" = "Yes Stupp\nUnmethylated")) + 
  scale_color_manual(values = c("no.M" = "#00AFBB", 
                                "yes.M" = "#E7B800", 
                                "no.UM" = "#FC4E07", 
                                "yes.UM" = "forestgreen"),
                     labels = c("No Stupp\nMethylated", 
                                "Yes Stupp\nMethylated", 
                                "No Stupp\nUnmethylated", 
                                "Yes Stupp\nUnmethylated")) + 
  labs(color = "Stupps/MGMT Status")
```

```{r}
library(dplyr)
library(broom)

broom::tidy(kruskal.test(overall_survival ~ stupps_mgmt, data = df_stupps))
```

### Median Age Analysis

```{r}
median_age <- median(df$diagnosis_age)

df$median_age_status <- cut(df$diagnosis_age,
                    breaks = c(0, 66, 100),  
                    labels = c("Below Median", "Equal to or Above Median"),  
                    right = FALSE)

km_median_age <- survfit2(Surv(overall_survival, status == 1) ~ median_age_status, data = df)
ggsurvfit(km_median_age) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival of Groups Above and Below Median Age"
    )
```

We see significance here, what about for dividing dataset into thirds?

```{r}
df$age_thirds <- cut(df$diagnosis_age, 
                     breaks = c(0, 52, 72, 100), 
                     labels = c("31-51", "52-71", "72-91"), 
                     right = FALSE)

km_median_age <- survfit2(Surv(overall_survival, status == 1) ~ age_thirds, data = df)
ggsurvfit(km_median_age) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival of Age Groups (Thirds)"
    )
```

And fourths? 

```{r}
df$age_fourths <- cut(df$diagnosis_age, 
                     breaks = c(0, 46, 61, 77, 100), 
                     labels = c("31-45", "46-60", "61-76", "77-91"), 
                     right = FALSE)

km_median_age <- survfit2(Surv(overall_survival, status == 1) ~ age_fourths, data = df)
ggsurvfit(km_median_age) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival of Age Groups (Fourths)"
    )
```

### Is temozolomide associated with longer survival?

```{r}
df$num_chemo <- factor(df$num_chemo, levels = c("0", "<1", "1", "<2", "2", "≥3"))
km <- survfit2(Surv(overall_survival, status == 1) ~ num_chemo, data = df)
ggsurvfit(km) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival of Patients by Number of Chemotherapy Cycles"
    )
```

Temozolomide makes a huge difference in overall survival. We can see that patients that received three or more cycles had the longest survival times. Below is a summary table of each chemo group. 

```{r}
chemo_freq <- data.frame(table(df$num_chemo))
colnames(chemo_freq) <- c("Number of Chemo Cycles", "Count")
chemo_freq$`Percentage of Patients` <- round(((chemo_freq$Count)/(sum(chemo_freq$Count))*100), 2)

df_chemo <- df %>% filter(!is.na(num_chemo)) %>% filter(!is.na(overall_survival))

mean_survival_chemo <- df_chemo %>%
  group_by(num_chemo) %>%
  summarise(mean_survival = mean(overall_survival))

colnames(mean_survival_chemo) <- c("Number of Chemo Cycles", "Mean Overall Survival")

chemo_stats <- merge(chemo_freq, mean_survival_chemo, by = "Number of Chemo Cycles", all = TRUE)
chemo_stats %>% arrange(`Number of Chemo Cycles`) %>% gt()

broom::tidy(kruskal.test(overall_survival ~ num_chemo, data = df_chemo))
```

### Is radiation associated with longer survival time?

```{r}
df$num_radiations <- factor(df$num_radiations, levels = c("0", "<1", "1", "<2", "2", "≥3"))
km <- survfit2(Surv(overall_survival, status == 1) ~ num_radiations, data = df)
ggsurvfit(km) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival of Patients by Number of Radiation Cycles"
    )
```

```{r}
rad_freq <- data.frame(table(df$num_radiations))
colnames(rad_freq) <- c("Number of Radiation Cycles", "Count")
rad_freq$`Percentage of Patients` <- round(((rad_freq$Count)/(sum(rad_freq$Count))*100), 2)

df_rad <- df %>% filter(!is.na(num_radiations)) %>% filter(!is.na(overall_survival))

mean_survival_rad <- df_rad %>%
  group_by(num_radiations) %>%
  summarise(mean_survival = mean(overall_survival))

colnames(mean_survival_rad) <- c("Number of Radiation Cycles", "Mean Overall Survival")

rad_stats <- merge(rad_freq, mean_survival_rad, by = "Number of Radiation Cycles", all = TRUE)
rad_stats %>% arrange(`Number of Radiation Cycles`) %>% gt()

broom::tidy(kruskal.test(overall_survival ~ num_radiations, data = df_rad))
```

### Mitotic Activity vs Ki67 Correlation 

KM Curve for the categorical parts of the mitotic activity increase. 

```{r}
# KM curve for mitotic activity 
micat_df <- df %>% filter(mitotic_activity_increase %in% c("0", "1to5", "6to10", ">10"))
micat_df$mitotic_activity_increase <- factor(micat_df$mitotic_activity_increase, levels = c("0", "1to5", "6to10", ">10"))

mai_count <- data.frame(table(micat_df$mitotic_activity_increase))
colnames(mai_count) <- c("Mitotic Index", "Count")
mai_count %>% gt()
```

Looking at Kaplan-Meier survival curves 

```{r}
micat_df_nz <- df %>% filter(mitotic_activity_increase %in% c("1to5", "6to10", ">10"))
micat_df_nz$mitotic_activity_increase <- factor(micat_df_nz$mitotic_activity_increase, levels = c("1to5", "6to10", ">10"))

mai_km <- survfit2(Surv(overall_survival, status == 1) ~ mitotic_activity_increase, data = micat_df)
ggsurvfit(mai_km) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival by Mitotic Index Category"
    )

mai_km <- survfit2(Surv(overall_survival, status == 1) ~ mitotic_activity_increase, data = micat_df_nz)
ggsurvfit(mai_km) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival by Mitotic Index (Excluding 0)"
    )

mai_km <- survfit2(Surv(overall_survival, status == 1) ~ Ki67_LI, data = micat_df_nz)
ggsurvfit(mai_km) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival by Ki67 Proliferation Index Category"
    )
```

Seeing if there is an association between mitotic activity and Ki67 proliferation index.

```{r}
micat_df$Ki67_LI <- factor(micat_df$Ki67_LI, levels = c("<15%", "15%", ">15%"))

broom::tidy(fisher.test(table(micat_df$mitotic_activity_increase, micat_df$Ki67_LI)))
```

There is a slight association between mitotic activity and Ki67.

##### Looking at specific values of mitotic activity & Ki67

```{r}
df_sm <- df %>% filter(!is.na(specific_mitotic))

unique(df_sm$specific_mitotic)
df_sm$specific_mitotic <- factor(df_sm$specific_mitotic, levels = c(
  "0", "1", "2", "2to3", "3", "3to4", "4", "4to5", "5", "5to6",
  "6", "7", "7to8", "8", "8to9", "9", "10", "at least 10", "11",
  "12", "13", "14", "15", "16", "19", "20", "21"
))
df$specific_mitotic <- factor(df$specific_mitotic, levels = c(
  "0", "1", "2", "2to3", "3", "3to4", "4", "4to5", "5", "5to6",
  "6", "7", "7to8", "8", "8to9", "9", "10", "at least 10", "11",
  "12", "13", "14", "15", "16", "19", "20", "21"
))

df_sm_filt <- df %>%
  filter(!grepl("to", specific_mitotic) & !grepl("at least ", specific_mitotic))


sm_km <- survfit2(Surv(overall_survival, status == 1) ~ specific_mitotic, data = df_sm_filt)
ggsurvfit(sm_km) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival by Specific Mitotic Index"
    )
```


### Seizures and DVT? 

#### Seizures at Presentation/During Course

```{r}
tbl0 <- data.frame(table(df$seizures_presentation))
colnames(tbl0) <- c("Seizures at Presentation?", "Count")
tbl0$`Percentage of Patients` <- round(((tbl0$Count)/(sum(tbl0$Count)))*100, 2)
tbl0 %>% gt()
```

```{r}
tbl1 <- data.frame(table(df$seizures_later))
colnames(tbl1) <- c("Seizures During Course?", "Count")
tbl1$`Percentage of Patients` <- round(((tbl1$Count)/(sum(tbl1$Count)))*100, 2)
tbl1 %>% gt()
```


```{r}
tbl01 <- data.frame(table(interaction(df$seizures_presentation, df$seizures_later)))
colnames(tbl01) <- c("Seizures at Presentation/Seizure During Course", "Count")
tbl01$`Percentage of Patients` <- round(((tbl01$Count)/(sum(tbl01$Count)))*100, 2)
tbl01 <- separate(tbl01, `Seizures at Presentation/Seizure During Course`, into = c("Seizures at Presentation", "Seizures During Course"), sep = "\\.")

tbl01_clean <- tbl01 %>%
  filter(`Seizures at Presentation` != "unclear" & `Seizures During Course` != "unclear")

df$seizure_interaction <- interaction(df$seizures_presentation, df$seizures_later)

df_seizure <- df %>% filter(!is.na(overall_survival)) %>% filter(!is.na(seizure_interaction))
df_seizure <- df_seizure %>% filter(!grepl("unclear", seizure_interaction))

mean_survival_seizure <- df_seizure %>%
  group_by(seizure_interaction) %>%
  summarise(mean_survival = mean(overall_survival))
mean_survival_seizure <- separate(mean_survival_seizure, seizure_interaction, into = c("Seizures at Presentation", "Seizures During Course"), sep = "\\.")
mean_survival_seizure_clean <- mean_survival_seizure %>%
  filter(`Seizures at Presentation` != "unclear" & `Seizures During Course` != "unclear")


merged_seizure <- merge(tbl01_clean, mean_survival_seizure_clean, by = c("Seizures at Presentation", "Seizures During Course"), how = "outer")

colnames(merged_seizure) <- c("Seizures at Presentation", "Seizures During Course", "Count", "Percentage of Patients", "Mean Survival")

merged_seizure %>% gt()

broom::tidy(kruskal.test(overall_survival ~ seizure_interaction, data = df_seizure))

km <- survfit2(Surv(overall_survival, status == 1) ~ seizure_interaction, data = df_seizure)
ggsurvfit(km) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival by Seizure at Presentation/During Course"
    )
```



```{r}
df_sei_dvt <- df %>% filter(seizures_presentation != "unclear" & seizures_later != "unclear") 

km <- survfit2(Surv(overall_survival, status == 1) ~ seizure_interaction, data = df_sei_dvt)
ggsurvfit(km) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Survival of Patients With Seizures At Presentation"
    )

km <- survfit2(Surv(overall_survival, status == 1) ~ seizures_later, data = df_sei_dvt)
ggsurvfit(km) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Survival of Patients With Seizures During Course"
    )
```

#### DVT/PE at Presentation/During Course

```{r}
tbl2 <- data.frame(table(df$dvt_pe_presentation))
colnames(tbl2) <- c("DVT/PE at Presentation?", "Count")
tbl2$`Percentage of Patients` <- round(((tbl2$Count)/(sum(tbl2$Count)))*100, 2)
tbl2 %>% gt()
```

```{r}
tbl3 <- data.frame(table(df$dvt_pe_during))
colnames(tbl3) <- c("DVT/PE During Course?", "Count")
tbl3$`Percentage of Patients` <- round(((tbl3$Count)/(sum(tbl3$Count)))*100, 2)
tbl3 %>% gt()
```

```{r}
tbl23 <- data.frame(table(interaction(df$dvt_pe_presentation, df$dvt_pe_during)))
colnames(tbl23) <- c("DVT/PE at Presentation/DVT/PE During Course", "Count")
tbl23$`Percentage of Patients` <- round(((tbl23$Count)/(sum(tbl23$Count)))*100, 2)
tbl23 <- separate(tbl23, `DVT/PE at Presentation/DVT/PE During Course`, into = c("DVT/PE at Presentation", "DVT/PE During Course"), sep = "\\.")

df$dvt_pe_interaction <- interaction(df$dvt_pe_presentation, df$dvt_pe_during)

df_dvtpe <- df %>% filter(!is.na(overall_survival)) %>% filter(!is.na(dvt_pe_interaction))

mean_survival_dvtpe <- df_dvtpe %>%
  group_by(dvt_pe_interaction) %>%
  summarise(mean_survival = mean(overall_survival))
mean_survival_dvtpe <- separate(mean_survival_dvtpe, dvt_pe_interaction, into = c("DVT/PE at Presentation", "DVT/PE During Course"), sep = "\\.")

merged_dvt_pe <- merge(tbl23, mean_survival_dvtpe, by = c("DVT/PE at Presentation", "DVT/PE During Course"), how = "outer")

colnames(merged_dvt_pe) <- c("Seizures at Presentation", "Seizures During Course", "Count", "Percentage of Patients", "Mean Survival")

merged_dvt_pe %>% gt()

broom::tidy(kruskal.test(overall_survival ~ dvt_pe_interaction, data = df_dvtpe))

km <- survfit2(Surv(overall_survival, status == 1) ~ dvt_pe_interaction, data = df_dvtpe)
ggsurvfit(km) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival by DVT/PE at Presentation/During Course"
    )
```


```{r}
km <- survfit2(Surv(overall_survival, status == 1) ~ dvt_pe_presentation, data = df_sei_dvt)
ggsurvfit(km) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Survival of Patients With DVT/PE at Presentation"
    )

km <- survfit2(Surv(overall_survival, status == 1) ~ dvt_pe_during, data = df_sei_dvt)
ggsurvfit(km) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Survival of Patients With DVT/PE During Course"
    )
```


Looking at DVT/PE History

```{r}
dvt_pe_hist <- data.frame(table(df$dvt_pe_history))
colnames(dvt_pe_hist) <- c("History of DVT/PE?", "Count")
dvt_pe_hist %>% gt()
```

### Time Between Symptoms & Diagnosis 

```{r}
tsd_df <- df %>%
  summarise(
    min_time = min(time_symptoms_diagnosis, na.rm = TRUE),
    max_time = max(time_symptoms_diagnosis, na.rm = TRUE),
    mean_time = mean(time_symptoms_diagnosis, na.rm = TRUE),
    median_time = median(time_symptoms_diagnosis, na.rm = TRUE),
  )

colnames(tsd_df) <- c("Min Time", "Max Time", "Mean Time", "Median Time")
tsd_df %>% gt()

ggplot(df, aes(x = time_symptoms_diagnosis)) +
  geom_density(fill = "skyblue") +
  labs(title = "Density Plot of Time From Onset of Symptoms to Diagnosis", x = "Time (Months)", y = "Density") +
  theme_minimal()
```
```{r}
df <- df %>% mutate(
  tsd_group = case_when(
    time_symptoms_diagnosis >= 0 & time_symptoms_diagnosis <= 0.5 ~ "≤ 0.5 Months",
    time_symptoms_diagnosis > 0.5 ~ "> 0.5 Months"
  )
)

km_fit <- survfit2(Surv(overall_survival, status == 1) ~ tsd_group, data = df)
ggsurvfit(km_fit) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival by Time from Onset of Symptoms to Diagnosis (Median)"
    )
```

```{r}
df <- df %>% mutate(
  tsd_group = case_when(
    time_symptoms_diagnosis >= 0 & time_symptoms_diagnosis <= 1.86 ~ "≤ 1.86",
    time_symptoms_diagnosis > 1.86 ~ "> 1.86"
  )
)

km_fit <- survfit2(Surv(overall_survival, status == 1) ~ tsd_group, data = df)
ggsurvfit(km_fit) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival of Patients by Time from Onset of Symptoms to Diagnosis (Mean)"
    )
```

```{r}
df <- df %>% mutate(
  tsd_group = case_when(
    time_symptoms_diagnosis >= 0 & time_symptoms_diagnosis <= 0.5 ~ "≤ 0.5 Months",
    time_symptoms_diagnosis > 0.5 ~ "> 0.5 Months"
  )
)

km_fit <- survfit2(Surv(overall_survival, status == 1) ~ tsd_group + received_stupp, data = df)
ggsurvfit(km_fit) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival of Patients by Time from Symptoms to Diagnosis + Stupp Status"
    )
```

#### Time from Surgery to Start of Radiation

```{r}
ttr_df <- df %>%
  summarise(
    min_time = min(time_surgery_radiation, na.rm = TRUE),
    max_time = max(time_surgery_radiation, na.rm = TRUE),
    mean_time = mean(time_surgery_radiation, na.rm = TRUE),
    median_time = median(time_surgery_radiation, na.rm = TRUE),
  )

colnames(ttr_df) <- c("Min Time", "Max Time", "Mean Time", "Median Time")
ttr_df %>% gt()

ggplot(df, aes(x = time_surgery_radiation)) +
  geom_density(fill = "skyblue") +
  labs(title = "Density Plot of Time From Diagosis to Radiation", x = "Time (Weeks)", y = "Density") +
  theme_minimal()

df <- df %>% mutate(
  ttr_group = case_when(
    time_surgery_radiation >= 0 & time_surgery_radiation <= 6 ~ "0-1.5 Months",
    time_surgery_radiation > 6 ~ "> 1.5 Month"
  )
)

km_fit <- survfit2(Surv(overall_survival, status == 1) ~ ttr_group, data = df)
ggsurvfit(km_fit) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    title = "Overall Survival of Patients by Time from Diagnosis to Start of Radiation"
    )
```
```{r}
km_fit <- survfit2(Surv(overall_survival, status == 1) ~ tsd_group + ttr_group, data = df)
ggsurvfit(km_fit) + 
  add_pvalue("annotation", size = 5) + 
  labs(
    x = "Time",
    y = "Survival Probability",
    title = "Overall Survival by Time from Onset of Symptoms to Diagnosis (TSD)\n+ Time from Diagnosis to Start of Radiation (TTR)"
  ) +
  scale_color_manual(
    values = c("#F8766D", "#7CAE00", "#00BFC4", "#C77CFF"), 
    labels = c(
      "> 0.5 Months TSD, > 1.5 Months TTR", 
      "> 0.5 Months TSD, 0-1.5 Months TTR", 
      "≤ 0.5 Months TSD, > 1.5 Months TTR", 
      "≤ 0.5 Months TSD, 0-1.5 Months TTR"
    )
  ) +
  theme_minimal()
```